### 2020-06-05 reviews/response.txt

We thank the reviewers for their exceptionally helpful and thorough
comments. Thorough reviews take time, interest, and expertise! Here,
we respond to reviewers' "Comments for Revision", marked by (C), and
due to limited space explain only a selection of detailed comments,
marked by (D). However, we have addressed all of the reviewers
comments in our revised paper either by making the requested change or
by providing additional explanation.

Our revision includes two substantive additions and numerous
improvements to explanations, graphics, and presentation. The first
substantive addition is a corroboration of our airflow implications on
temperature with temperature data from another publication. The second
substantive addition is a new result that job scheduling and folded
torus interconnect were contributing factors to temperature patterns,
and explain a peculiar interleaving column hazard variability.

We have added more explantion to most figures, making them more
self-contained and relevant to our conclusions.

@@

We now have location-specific information for average temperature
across cages and it agrees with our airflow effect on temperature
interpretation (V.-6th paragraph).

(D)

Reviewer 2:

(C)
We have strengthened our explanations in several sections and
particularly around our airflow explanation in Fig.4. Both substantive
improvements described above relate to this point, on one hand
strengthening it with temperature data, and on the other hand also finding
a more nuanced impact of scheduling.

(D)
Section 3: see changes in first paragraph.

Reviewer 3:

(C)
Generalization: This analysis is already generating interest from
Summit operations. A lesson learned is that Titan was not well
instrumented for data collection - thus our difficulty in obtaining
temperature data. Frontier will likely have even more
instrumentation.

Novelty in graphics: Figures 1 and 5 were specifically designed for
this data using R's amazingly high-level and flexible ggplot2
concepts. Figures 10-13 are standard in R's survival analysis
packages.

Reviewer 4:

(C)

### 2020-06-05 dataclean.tex
some rephrasing

As most analysis software relies on rectangular table-like data, we
fill the needed repeats of values missing in the raw data (see
Fig.~\ref{fig:dataraw}). To focus our analysis on data after the first
two rework cycles in the break-in period, we first reduce the data to
the GPUs that were installed shortly before 2014 started. This was
done by removing data for any units with a {\em remove} date before
2014. After this reduction, there were still six older units
remaining, which we also removed to have a clean set of units for the
analysis. We do some further processing to handle time overlaps in a
tiny fraction (under 0.0007 in GPU life and 0.0002 in location life)
of recorded life in the raw records by simply dropping the overlapping
life times. \rev{A detailed followup of a few overlaps found that they
  are caused by incomplete inventories, where a node query times out,
  creating the appearance that a unit is in two locations.}

>>>>

As most analysis software relies on rectangular table-like data, we
fill the needed repeats of values missing in the raw data (see
Fig.~\ref{fig:dataraw}). To focus our analysis on data after the first
two rework cycles in the break-in period, we first reduce the data to
the GPUs that were installed shortly before 2014 started. This was
done by removing data for any units with a {\em remove} date before
2014. After this reduction, there were still six older units
remaining, which we also removed to have a clean set of units for the
analysis. We do some further processing to handle time overlaps in a
tiny fraction (under 0.0007 in GPU life and 0.0002 in location life)
of recorded life in the raw records by simply dropping the overlapping
life times. \rev{A detailed followup of a few of these overlaps found
  that they are caused by incomplete inventories, where for a moved
  GPU a node query of times out, creating the appearance that a unit
  is in two locations.}

### 2020-06-05 discussion.tex
Renamed the Discussion Section to Conclusions.
some rephrasing

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Discussion
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
\label{section:discussion}

The experienced unexpected reliability issues with the Titan supercomputer raise
several questions and discussion points. First, there is the mitigation component
that keeps the system operating at an acceptable level and possibly even restores
some, if not all, of the lost capability. Second, there is the long-term planning
component that addresses such scenarios in future-generation supercomputers.

>>>>

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Conclusions
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\rev{Conclusions}}
\label{section:discussion}

\rev{The failure rates of the old batch are not matched by the new
  batch nor by experience at other facilities with the same type of
  GPU components, making this an unexpected event that had
  considerable impact on operations and on availability of the
  system. On a positive but ironic note, the corrosion process in the
  failing resistors made the GPUs act as a sensitive instrument for
  obtaining long-term integrals of component heat dynamics and
  provided lessons learned. Here we dicsuss the conclusions we can
  draw from the experience on mitigation, which keeps the system
  operating at an acceptable level and possibly even restores some of
  the lost capability, and on long-term planning, which addresses such
  scenarios in future-generation supercomputers.}
